---
title: "Bayesian Analysis of Likert Categorical Responses"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

Survey data typically come in the form of likert 5 and 7 point scales and are categorical in nature. This limits analysis to $\chi^2$ analysis, Fischer's exact test and conversion of the categorical scales to numeric values and a t-test analysis. The former doesn't lead to understanding of pairwise differences and the latter isn't faithful to the categorical nature of the variables.

# Proposal

Use Bayesian analysis and count data in order to understand the differences in the data.

# Method

Generate some fake survey data, pass through a Binomial liklihood with a Beta prior. Summarise the posteriors to understand the differences.

## Fake Data

```{r fake_data}
set.seed(336)

questions <- factor(
  c("Strongly agree", "Agree", "Neither", "Disagree", "Strongly Disagree"),
  c("Strongly Disagree", "Disagree", "Neither", "Agree", "Strongly agree"))

# Establish the vectors of probabilties
p1 <- c(.25, .25, .05, .30, .15)
p2 <-c(.25, .20, .02, .33, .20)

group_1 <- rbinom(5, 100, p1)+rbinom(5,5,.25)
group_2 <- rbinom(5, 100, p2)++rbinom(5,5,.25)

df <- data.frame(questions, group_1, group_2)

knitr::kable(df)
```

## Frequentist method

For comparison:

```{r}
chisq.test(df$group_1, df$group_2)
```

Results are not significant, these two distributions are drawn from the same distribution...so the frequentist analysis indicates.

## Posterior Draws

The Bayesian Analysis will take the following form:

Liklihood:

$$Y|\theta \sim Binomial(N, \theta)$$
Prior
$$\theta \sim Beta(a,b)$$
Posterior:
$$\theta | Y \sim Beta(Y +a, N- Y +b)$$
Where a = 0.1, b = 0.1, Y is number of positive reponses for the question and N is the total number of responses.

Set up for the Bayesian analysis

```{r}
#Prep for MC
group_1_n <- sum(group_1)
group_2_n <- sum(group_2)

#Priors
a <- .1
b <- .1

#Iterations
sampz <- 100000

#Draw from posterior
group_1_posterior <-list()
for(i in 1:length(group_1)){
  group_1_posterior[[i]] <- rbeta(sampz,group_1[i]+a,group_1_n-group_1[i]+b)
  
}

group_2_posterior <-list()
for(i in 1:length(group_2)){
  group_2_posterior[[i]] <- rbeta(sampz,group_2[i]+a,group_2_n-group_2[i]+b)
  
}
```

## Examine Posterior Distribution

```{r echo=FALSE}
par(mfrow = c(5,2), mar = c(.1,.1,.1,.1))
hist(group_1_posterior[[1]], breaks =60, main = "Group 1-Strongly agree", 
     xlim = c(0,.5))
abline(v = mean(group_1_posterior[[1]]), col = "red")
hist(group_2_posterior[[1]], breaks =60, main = "Group 2-Strongly agree", 
     xlim = c(0,.5))
abline(v = mean(group_2_posterior[[1]]), col = "red")

hist(group_1_posterior[[2]], breaks =60, main = "Group 1-Agree", 
     xlim = c(0,.5))
abline(v = mean(group_1_posterior[[2]]), col = "red")
hist(group_2_posterior[[2]], breaks =60, main = "Group 2-Agree", 
     xlim = c(0,.5))
abline(v = mean(group_2_posterior[[2]]), col = "red")

hist(group_1_posterior[[3]], breaks =60, main = "Group 1-Neither", 
     xlim = c(0,.5))
abline(v = mean(group_1_posterior[[3]]), col = "red")
hist(group_2_posterior[[3]], breaks =60, main = "Group 2-Neither", 
     xlim = c(0,.5))
abline(v = mean(group_2_posterior[[3]]), col = "red")

hist(group_1_posterior[[4]], breaks =60, main = "Group 1-Disagree", 
     xlim = c(0,.5))
abline(v = mean(group_1_posterior[[4]]), col = "red")
hist(group_2_posterior[[4]], breaks =60, main = "Group 2-Disagree", 
     xlim = c(0,.5))
abline(v = mean(group_2_posterior[[4]]), col = "red")

hist(group_1_posterior[[5]], breaks =60, main = "Group 1-Strongly Disagree", 
     xlim = c(0,.5))
abline(v = mean(group_1_posterior[[5]]), col = "red")
hist(group_2_posterior[[5]], breaks =60, main = "Group 2-Strongly Disagree", 
     xlim = c(0,.5))
abline(v = mean(group_2_posterior[[5]]), col = "red")


```

```{r echo=FALSE}
suppressPackageStartupMessages(library(tidyverse))
#name my lists
try_1<-list()
for(i in 1:5){
  catz <- rep(as.character(questions[i]),sampz)
  try_1[[i]] <- group_1_posterior[[i]] %>%
    as_data_frame() %>%
    add_column(., catz)
}

try_2<-list()
for(i in 1:5){
  catz <- rep(as.character(questions[i]),sampz)
  try_2[[i]] <- group_2_posterior[[i]] %>%
    as_data_frame() %>%
    add_column(., catz)
}
```


```{r echo=FALSE}
library(ggplot2)
library(dplyr)
try_1 %>%
  bind_rows() %>%
  mutate(group = "group_1") %>%
  union(., try_2 %>%
  bind_rows() %>% 
  mutate(group = "group_2")) %>%
  mutate(catz = factor(catz, c("Strongly Disagree", "Disagree", "Neither", "Agree", "Strongly agree"))) %>% 
  ggplot(aes(value, color = group, group = group))+
  geom_density()+
  facet_wrap(~catz)+theme_minimal()
```


## Summarise Results

First, I need a helper function to summarise the two different groups.

```{r}
#Helper Function
library(purrr)

summarise_posterior <- function(x,y){
  mu_1 <- map_dbl(x, mean)
  mu_2 <- map_dbl(y, mean)
  
  sd_1 <- map_dbl(x, sd)
  sd_2 <- map_dbl(y, sd)
  
  delta_<-list()
  pooled_sd <- list()
  pro_difference <-vector()
  for(i in 1:length(x)){
    delta_[[i]] <- x[[i]]-y[[i]]
    delta_mu <- map_dbl(delta_, mean)
    pooled_sd[[i]] <-sqrt(var(x[[i]]) + var(y[[i]]))
    cohens <-delta_mu/ unlist(pooled_sd)
    pro_difference[i] <- max(mean(x[[i]] > y[[i]]), 
                          1-mean(x[[i]] > y[[i]]))
  }
  
  out <- cbind(mu_1, mu_2, sd_1, sd_2,delta_mu, cohens,pro_difference)
  
  out
}

```


Now look at the results in table form.

```{r echo=FALSE}
statz <- summarise_posterior(group_1_posterior, group_2_posterior)

knitr::kable(data.frame(question =as.character(questions),statz, p1, p2 ), digits = 3)
```

# Summary

Now we can look at the individual differences as well as the individual propabilties of difference.

# Power Analysis

```{r}

group_1_success <- 1

group_2_success <- 5


a_A <- group_1_success+1
b_A <- group_1_n-group_1_success+1
a_B <- group_2_success+1
b_B <- group_2_n-group_2_success+1

fun <- function(i) exp(lbeta(a_A+i, b_B+b_A)
             - log(b_B+i)
             - lbeta(1+i, b_B)
             - lbeta(a_A, b_A))

sum(vapply(0:(a_B-1), fun, numeric(1)))
```


#Frequentist Power Calculations

```{r}
p <- 0.25 #Treatment Group
p0 <- 0.15
alpha <- 0.1 # Type
beta <- 0.2 # Power
(n=p*(1-p)*((qnorm(1-alpha/2)+qnorm(1-beta))/(p-p0))^2)
ceiling(n) # 50
z=(p-p0)/sqrt(p*(1-p)/n)
(Power=pnorm(z-qnorm(1-alpha/2))+pnorm(-z-qnorm(1-alpha/2)))
```

